Weather events impact in United States injuries and deaths (1950-2011) 
========================================================

## Synopsis
This data come from the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database. The objective was to analyze the effects of different types of natural disasters over the outcomes (fatalities and injuries) related to the people of the U.S.

### Data Processing
This database is stored as a comma-separated-value file compressed via the bzip2 algorithm to reduce its size. You can download the file from this link: [Storm Data 47Mb](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2)

You can read the file by extracting into you working directory and using the function read.csv.
I strongly recommend that you format the column BGN_DATE with the functions as.POSIXlt for any further investigation related to month/year of this data. However it is not needed for this short analysis.
The graphics of this work are based on the ggplot2 codes. So for proper working use the library function to load ggplot2 before the analysis.


```{r}
StormData <- read.csv("repdata-data-StormData.csv")
df <- within(StormData, datetime <- as.POSIXlt(BGN_DATE, format="%m/%d/%Y"))
library(ggplot2)
require(gridExtra)
```

A few tables ware constructed. The first table, called type, was intend to check the number of different disasters and its frequency. This same result can be seen by the use of the function summary, however, this table will be used for later propose. With the use of this last function, we can also see the mean fatalities and injuries, and the most common state to have natural disasters.
The use of the function str shows us that there are 985 different natural disasters.

```{r}
type <- data.frame(table(StormData$EVTYPE))
summary(df[,c(7,8,23,24)],10)
str(type)

```

Other tables were made to sum the number of injuries or deaths for each type of disasters.
Though the column bind function, these sums were attach to the "type" table. From this combined table it was possible to check the mean values of fatalities and injuries for each type of disaster. These tables were ordered by most frequent.
A table of the separated number of outcomes order by the most "chaotic" was also created.

```{r}

dffatsum <- aggregate(FATALITIES ~ EVTYPE, data = df, sum, na.action = na.omit)
dfinjsum <- aggregate(INJURIES ~ EVTYPE, data = df, sum, na.action = na.omit)

dffs <- cbind(type, dffatsum$FATALITIES, dfinjsum$INJURIES)
colnames(dffs) <- c("EVTYPE", "FREQ", "FATALITIES", "INJURIES")
dffs$FATMEAN <- dffs$FATALITIES / dffs$FREQ
dffs$INJMEAN <- dffs$INJURIES / dffs$FREQ

FATALITIES <- head(dffs[order(dffs$FATALITIES, decreasing=TRUE),],20)[,c(1,3)]
INJURIES <- head(dffs[order(dffs$INJURIES, decreasing=TRUE),],20)[,c(1,4)]

FATMEAN <- head(dffs[order(dffs$FATMEAN, decreasing=TRUE),],20)[,c(1,5)]
INJMEAN <- head(dffs[order(dffs$INJMEAN, decreasing=TRUE),],20)[,c(1,6)]

dffat <- head(df[order(df$FATALITIES, decreasing=TRUE),],40)[,c(8,23)]
dfinj <- head(df[order(df$INJURIES, decreasing=TRUE),],70)[,c(8,24)]

```

### Results
For exception of tornados, it is possible to see that the most 10 frequent catastrophe are not as well the most responsible for the highest number of outcomes related to people.
Bellow it's possible to show the 10 most frequent disasters:


```{r}
head(dffs[order(dffs$FREQ, decreasing=TRUE),],10)

```

In addition, due to the relationship of frequency of events, the disasters with the highest number of incidents doesn't have the highest mean (outcome/ disaster frequency). These numbers shows us that the average number of deaths and injuries doesn't represent a valid analysis, since the mean is highest to rare events. However, this brings the knowledge that people are less prepared to these events, therefore, they bring higher risks of incidents.
```{r}
head(dffs[order(dffs$FATMEAN, decreasing=TRUE),],10)
head(dffs[order(dffs$INJMEAN, decreasing=TRUE),],10)

```
A graphical interpretation of the 20 events with highest average:
```{r fig.width=11}
plot3 <- ggplot(FATMEAN, aes(as.character(EVTYPE), FATMEAN)) + geom_bar(stat = "identity", 
        position = "dodge") + theme(axis.text.x = element_text(angle = 90)) + 
        labs(title = "Top 20 disasters with most number of deaths per event", 
        x = NULL, y = "Number of Fatalities")
plot4 <- ggplot(INJMEAN, aes(as.character(EVTYPE), INJMEAN)) + geom_bar(stat = "identity", 
        position = "dodge") + theme(axis.text.x = element_text(angle = 90)) + 
        labs(title = "Top 20 disasters with most number of injuries per event", 
        x = NULL, y = "Number of Injuries")
grid.arrange(plot3, plot4, ncol = 2)
```

Furthermore, to better interpretation of this data, an analysis of the sum of outcomes can be better analyzed with a graphical view. This graphic shows that tornados are the biggest responsible for injuries and fatalities even though its average is low.
```{r fig.width=11}
plot1 <- ggplot(FATALITIES, aes(as.character(EVTYPE), FATALITIES)) + geom_bar(stat = "identity", 
        position = "dodge") + theme(axis.text.x = element_text(angle = 90)) + 
        labs(title = "Top 20 disasters with most number of deaths", 
        x = NULL, y = "Number of Fatalities")
plot2 <- ggplot(INJURIES, aes(as.character(EVTYPE), INJURIES)) + geom_bar(stat = "identity", 
        position = "dodge") + theme(axis.text.x = element_text(angle = 90)) + 
        labs(title = "Top 20 disasters with most number of injuries", 
        x = NULL, y = "Number of Injuries")
grid.arrange(plot1, plot2, ncol = 2)

subset(dffs, dffs$EVTYPE == "TORNADO")
```

In the next graphic, it's possible to show the 7 worst and most devastating event of each disasters.
By this graphic we can see that, even though tornados are the most frequent cause of deaths and injuries overall, excessive heat, extreme heat and heat, when happens, have the highest impact in numbers of fatalities.

```{r fig.width=11}
plot5 <- ggplot(dffat, aes(as.character(EVTYPE), FATALITIES)) + geom_bar(stat = "identity", 
        position = "dodge") + theme(axis.text.x = element_text(angle = 90)) + 
        labs(title = NULL, 
        x = NULL, y = "Number of Fatalities")
plot6 <- ggplot(dfinj, aes(as.character(EVTYPE), INJURIES)) + geom_bar(stat = "identity", 
        position = "dodge") + theme(axis.text.x = element_text(angle = 90)) + 
        labs(title = NULL, 
        x = NULL, y = "Number of Injuries")
grid.arrange(plot5, plot6, ncol = 2, main = "Most devastating disater for the worst 7 types of events")
```

