Practical Machine Learning Assignment
========================================================

The data consists of a set of readings from sensors attached to subjects performing a variety of exercises. The task is to build a model which will correctly classify activities, given unlabelled sets of readings.

Load the [Caret](http://caret.r-forge.r-project.org/) library and set the random number generator's seed, to ensure reproducibility:

```{r setup, message=FALSE}
library(caret)
set.seed(1234)
```

## Reading and Cleaning the Data
```{r}
rawData <- read.csv("pml-training.csv", na.strings=c("NA",""), strip.white=T)
dim(rawData)
na <- apply(rawData, 2, function(x) { sum(is.na(x)) })
cleandata <- subset(rawData[, which(na == 0)], select=-c(X, user_name, new_window, num_window, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp))
dim(cleandata)

```

### Partition
Partitioning the data into training and test 

```{r}
inTrain <- createDataPartition(cleandata$classe, p=0.7, list=F)
training <- cleandata[inTrain,]
testing <- cleandata[-inTrain,]
```

## Training a Random Forest Model

Now use the train model on the training set.

```{r, cache=TRUE, message=FALSE}
ctrl <- trainControl(allowParallel=T, method="cv", number=4)
rawdata2 <- read.csv("pml-testing.csv", na.strings=c("NA",""), strip.white=T)
testdata <- subset(rawdata2[, which(na == 0)], 
                        select=-c(X, user_name, new_window, num_window, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp))

validdata <- subset(cleandata, select=c(roll_belt, pitch_forearm, yaw_belt, magnet_dumbbell_y, pitch_belt, magnet_dumbbell_z, roll_forearm, accel_dumbbell_y, roll_dumbbell, magnet_dumbbell_x,classe))
model <- train(classe ~ ., data=validdata[inTrain,], model="rf", trControl=ctrl)



```

## Predictions
```{r}
predict(model, newdata=testdata)
pred <- predict(model, newdata=testing)
sum(pred == testing$classe) / length(pred)
confusionMatrix(testing$classe, pred)$table
```


I use this method because it is fast and its accuracy on the test set is 98.5%.

